The state of knowledge for marine invertebrate biodiversity in the continental US

### 2017-09-25 manuscript.Rmd
small edits

Research attempting to understand how much we know about marine biodiversity has focused on estimating the total number of species that inhabits the oceans, and which fraction of this diversity has been named [@Appeltans2012,@Costello2013]. Undoubtedly, the development of molecular methods to investigate species limits has reinvigorated taxonomy, and contributed to elucidating cryptic species complexes [@Knowlton2000]. Thus, it has been argued that molecular data can speed up the enterprise of discovering and naming species [@Blaxter2016]. As species remain the most convenient unit to measure biological diversity, understanding how much has been done, how much remains to be done, and how we can speed up the description of species diversity is crucial. However, to understand the role of these species in their environment, we need to document the ecology of these species [@Cernansky2017]. For instance, to be able to predict how their geographic ranges will be affected by global change, we need detailed information about their environmental requirements. Occurrence data for marine species could be used for environmental niche modeling of their geographic and bathymetric distributions but the proportion of species for which we have enough data to do so is unknown.

Until very recently, our knowledge of biodiversity was scattered among heterogeneous sources (e.g. natural history collections, publications, expert knowledge), difficult to access, and therefore difficult to synthesize. The development of public databases for specimen records (e.g. iDigBio [@Page2015]), species occurrences (e.g., OBIS [@Zhang2002]), genetic data (e.g., GenBank [@Benson2013], BOLD [@Ratnasingham2007a]), and registers of species names (e.g., WoRMS [@Costello2013b]), now allow us to evaluate what and how much we know. Here, we aggregate publicly available data for marine invertebrates that occur within the limits of the exclusive economic zone (EEZ) of the continental United States (US), as well as the Gulf of Mexico (GOM) and the Pacific Northwest (PNW), to highlight gaps in our knowledge that provide a road map for important new directions in global research. We show that while public databases are capturing a significant portion of our knowledge about species richness, most species are known from only a few specimens and a few locations. We emphasize that our understanding of marine biodiversity, and the impact of human activities on biological communities require not only a more accurate list of the species that inhabits the oceans, but also more comprehensive data on species distributions and abundances. These goals can only be achieved with additional cost-effective sampling, and a commitment to training more taxonomists and experts that can identify and interpret the material collected.

@@

Missing geographic coordinates reduces the overlap between the list of species from the regional list and the databases. Using iDigBio we found that for the `r format(not_really_in_database$idig_gom_n_spp_not_in_db, big.mark=",")` species that do not appear within the geographic range of the Gulf of Mexico but are on the list, `r format_output(not_really_in_database$idig_gom_p_spp_not_in_db*100)`% are completely absent from iDigBio, while for the `r format(not_really_in_database$idig_gom_n_spp_no_coords, big.mark=",")` species that do appear in iDigBio, `r format_output(not_really_in_database$idig_gom_p_spp_no_coords*100)`%  do not have geographic coordinates associated with them, and for the ones that do have geographic coordinates, `r format_output(not_really_in_database$idig_gom_p_spp_within_usa*100)`% ($n=$  `r format(not_really_in_database$idig_gom_n_spp_within_usa, big.mark=",")`) fall within US waters. Additional digitization of collections, including the capture of their geographic coordinates information, are needed to improve biodiversity knowledge at regional scales. For instance, one can expect that the digitization coverage for the Gulf of Mexico will greatly increase once records from the Florida Wildlife Research Institute collections (St Petersburg, FL) become integrated into biodiversity information systems.

@@

The above estimates are based on species names, but many traditional species include multiple cryptic species [@Knowlton2000;@Meyer2005-evolution]. For example, up to `r format_output(100 * stat_bold$p_spp_multi_bin)`% of `r format(stat_bold$n_spp_not_singleton, big.mark=",")` species represented by more than 1 record in BOLD include cryptic species, as denoted by assignment to  more than 1 BIN (BOLD Index Number, operational taxonomic units that closely resemble species [@Ratnasingham2013]). Specimens identified under the same species but sampled from distant locations, are especially prone to representing cryptic species. Geographic information is missing for most records (`r format_output(100*stat_bold$p_no_coords)`% of the `r format(stat_bold$n_total_records, big.mark=",")` BOLD records analyzed here have no geographic coordinate information associated with them). It is thus crucial to ensure that accurate geographic information is provided when DNA barcodes are published, and that enough samples for each species are sequenced, to capture more accurately their geographic range and the extent of their genetic diversity.

@@

With data aggregation of natural history collections through initiatives like iDigBio, and the compilation of a nearly complete register of marine species with WoRMS, in combination with the development of algorithms for text analysis, it is increasingly easier to validate and spell check the scientific names of the identifications of digitized museum specimens. Linking species names to identifiers in taxonomic databases would facilitate data linking, specimen discoverability, and data usability. Additionally, because species names registers like WoRMS link the history of the names and their synonymies, it is possible to dissociate the concepts of species identification with the concept of species identity, allowing researchers to compare specimens identified at different time periods.

>>>>

Research attempting to understand how much we know about marine biodiversity has focused on estimating the total number of species that inhabits the oceans, and which fraction of this diversity has been named [@Appeltans2012,@Costello2013]. Undoubtedly, the development of molecular methods to investigate species limits has reinvigorated taxonomy, and contributed to elucidating cryptic species complexes [@Knowlton2000]. Thus, it has been argued that molecular data can speed up the enterprise of discovering and naming species [@Blaxter2016]. As species remain the most convenient unit to measure biological diversity, understanding how much has been done, how much remains to be done, and how we can speed up the description of species diversity is crucial. However, to understand the role of these species in their environment, we need to document their ecology [@Cernansky2017]. For instance, to be able to predict how their geographic ranges will be affected by global change, we need detailed information about their environmental requirements. Occurrence data for marine species could be used for environmental niche modeling of their geographic and bathymetric distributions but the proportion of species for which we have enough data to do so is unknown.

Until very recently, our knowledge of biodiversity was scattered among heterogeneous sources (e.g. natural history collections, publications, expert knowledge), difficult to access, and therefore difficult to synthesize. The development of public databases for specimen records (e.g. iDigBio [@Page2015]), species occurrences (e.g., OBIS [@Zhang2002]), genetic data (e.g., GenBank [@Benson2013], BOLD [@Ratnasingham2007a]), and registers of species names (e.g., WoRMS [@Costello2013b]), now allow us to evaluate what and how much we know. However, few studies have sought to synthesize existing knowledge by combining information from these databases.  Here, we aggregate publicly available data for marine invertebrates that occur within the limits of the exclusive economic zone (EEZ) of the continental United States (US), as well as the Gulf of Mexico (GOM) and the Pacific Northwest (PNW), to highlight gaps in our knowledge that provide a road map for important new directions in global research. We show that while public databases are capturing a significant portion of our knowledge about species richness, most species are known from only a few specimens and a few locations. We emphasize that our understanding of marine biodiversity, and the impact of human activities on biological communities require not only a more accurate list of the species that inhabits the oceans, but also more comprehensive data on species distributions and abundances. These goals can only be achieved with additional cost-effective sampling, and a commitment to training more taxonomists and experts that can identify and interpret the material collected.

@@

Missing geographic coordinates reduces the overlap between the list of species from the regional list and the databases. Using iDigBio, we found that for the `r format(not_really_in_database$idig_gom_n_spp_not_in_db, big.mark=",")` species that do not appear within the geographic range of the Gulf of Mexico but are on the list, `r format_output(not_really_in_database$idig_gom_p_spp_not_in_db*100)`% are completely absent from iDigBio, while for the `r format(not_really_in_database$idig_gom_n_spp_no_coords, big.mark=",")` species that do appear in iDigBio, `r format_output(not_really_in_database$idig_gom_p_spp_no_coords*100)`%  do not have geographic coordinates associated with them, and for the ones that do have geographic coordinates, `r format_output(not_really_in_database$idig_gom_p_spp_within_usa*100)`% ($n=$  `r format(not_really_in_database$idig_gom_n_spp_within_usa, big.mark=",")`) fall within US waters. Additional digitization of collections, including the capture of their geographic coordinates information, are needed to improve biodiversity knowledge at regional scales. For instance, one can expect that the digitization coverage for the Gulf of Mexico will greatly increase once records from the Florida Wildlife Research Institute collections (St Petersburg, FL) become integrated into biodiversity information systems.

@@

The above estimates are based on species names, but many traditional species include multiple cryptic species [@Knowlton2000;@Meyer2005-evolution]. For example, up to `r format_output(100 * stat_bold$p_spp_multi_bin)`% of `r format(stat_bold$n_spp_not_singleton, big.mark=",")` species represented by more than 1 record in BOLD include cryptic species, as denoted by assignment to  more than 1 BIN (BOLD Index Number, operational taxonomic units that closely resemble species [@Ratnasingham2013]). Specimens identified under the same species but sampled from distant locations, are especially prone to representing cryptic species. Geographic information is missing for most records (`r format_output(100*stat_bold$p_no_coords)`% of the `r format(stat_bold$n_total_records, big.mark=",")` BOLD records analyzed here have no geographic coordinate information associated with them). It is thus crucial to ensure that precise and accurate geographic information is provided when DNA barcodes are published, and that enough samples for each species are sequenced, to capture more accurately their geographic range and the extent of their genetic diversity.

@@

With data aggregation of natural history collections through initiatives like iDigBio, and the compilation of a nearly complete register of marine species with WoRMS, in combination with the development of algorithms for text analysis, it is increasingly easier to validate and spell check the scientific names of the identifications of digitized museum specimens. Linking species names to identifiers in taxonomic databases would facilitate data linking, specimen discoverability, and data usability. Additionally, because species name registers like WoRMS link the history of the names and their synonymies, it is possible to dissociate the concepts of species identification with the concept of species identity, allowing researchers to compare specimens identified at different time periods.
